Joint training of classification model and similarity model for low-resource text classification in chatbot

### 2020-08-10 chatbot_classification/aaai_2021/paper.tex
add subsections in Result Analysis

\begin{figure}[htbp]
\centering 
\subfigure[classification model]{
\label{Fig.1.sub.1}
\includegraphics[scale=0.7]{picture1.pdf}}
\subfigure[similarity model]{
\label{Fig.1.sub.2}
\includegraphics[scale=0.55]{picture2.pdf}}
\caption{Two popular structures of building task-specific conversational chatbot}
\label{Fig.1.main}
\end{figure}

@@

\begin{figure}[htbp]
\begin{centering}
\includegraphics[scale=0.48]{picture3}
\par\end{centering}
\caption{\textbf{Network Structure of 2-stage SFC:} a joint system of multi-task for both text classification task and sentence-pair semantic similarity task.}
\label{fig:ver.1.3}
\end{figure}

@@

\begin{figure*}[t]
\begin{centering}
\includegraphics[scale=0.72]{picture4} 
\par\end{centering}
\caption{\textbf{Network Structure of hierarchical SFC:} a joint model of sentence-pair model and text classification model organized in hierarchical structure by completing different tasks at different network layers.}
\label{fig:framework}
\end{figure*}

@@

$\bullet$ \textbf{full-scale real-world task-specific faq chatbot dataset}

@@

$\bullet$ \textbf{subset of customer review dataset for extreme classification}

@@

$\bullet$ \textbf{subset of recently-released intent detection datasets}.

@@

First, we turn to compare among the four SFC variants. We conduct ablation experiments to study the benefits brought by multi-task training by comparing multi-task 2-stage SFC with single-task 2-stage SFC. We can see from Table \ref{tbe:table2} that 2-stage SFC with only task 1 (sentence-pair similarity task) degrades by 0.65 percent on average and 2-stage SFC with only task 2 (top-K classification task) degrades by 0.3 percent on average. The degradation indicates that training with these 2 tasks together in SFC can effectively help the sentence-pair model extract more knowledge from the limited amount of data. Besides, task 2 degrades 0.35 percent less than task 1. This observation suggests that task 2 may play a relatively more important role in multi-task learning, which leads us to put larger weight $\alpha_{K}$ to the loss $\mathcal{L}^{K}$ of task 2 when calculating the total loss $\mathcal{L}$. Also, we can see that the hierachical SFC steadily outperforms 2-stage SFC on 5 diverse datasets by 2 percents on average. This observation supports the idea that the hierachical joint model structure can alleviate the limitation brought by the lack of interaction between similarity model and classification model in 2-stage SFC. Therefore, for the following analysis, we will focus on the comparison between hierachical SFC and the other baseline models.

Next, we analyze the overall improvement trend of SFC under various few-shot settings. Our main focus is to study chatbot building in common situation during real-world application where only a few sample sentences are available for each class. The experimental results for the 3 diverse intent detection datasets under few-shot settings indicate that our proposed SFC can achieve +2.47 percents average improvement over one of the most powerful baseline models, which is RoBERTa-large classification model, on all these few-shot settings. Moreover, the improvement in F-score becomes more and more prominent with the decrease of data sample number per class. Our hierarchical SFC can achieve +4.97 percents average improvements over RoBERTa-large classification model in the most extreme few-shot setting with only 5 training samples per class.

Then, we begin to compare the general performance of SFC with classification model and similarity model baselines. As for the classification model, hierarchical SFC outperforms BERT-large by +9.81 percents, ALBERT-xxlarge by 4 percents, RoBERTa-large by +2.47 percents, RoBERTa-base by +4.52 percents on average F-score for 5 datasets. In this way, we can see that RoBERTa is the best pre-trained model for the low-resource text classification task we are studying. As for the similarity model, we only choose the strongest pretrained model RoBERTa-large as our baseline, and hierarchical SFC achieves the improvement of 3 percents on average. Moreover, hierarchical SFC's result on 2 datasets without few-shot settings, which are ITG and Amazon-670k, are also 1.24 percents and 2.89 percents higher than Roberta-large classification model. These improvements illustrate that putting classification model and similarity model at different network layers to cooperate with each other is a successful design of hierarchical network structure for joint model, which enables these two models assist each other for better overall results.

Finally, we analyze the effect brought by different settings of hyperparameters, see Table \ref{tbe:table3}. We conduct experiments on 5 different settings of hyperparameters, which are the number of candidate class number $K$ and the number of sentence pairs $P$ for each class, on ITG and Amazon-670K datasets to find the most suitable setting of these 2 hyperparameters. We control the total number of sentence pairs sampled for each user query into about the same level of amount(50-60). Through analyzing the experimenal results shown in Table \ref{tbe:table3}, we find that the best setting of hyperparameters might be slightly different for datasets of different data size. The class label number of ITG and Amazon-670k are similar in our experimental setting, which is around 250. However the total amount of data points of ITG is approximately 2 times larger than Amazon-670k. Therefore, the setting of $K=5$ and $P=10$ is good enough for ITG since most of the classes contains over 10 data points. However, for Amazon-670k, the amount of data sample for each class is scarce, thus, we may need more candidate classes to sample enough effective sentence pairs for sentence-pair model layer of SFC. In this way, the setting of $K=10$ and $P=5$ becomes a good choice.

>>>>

\begin{figure}[htbp]
\centering 
\subfigure[classification model]{
\label{Fig.1.sub.1}
\includegraphics[scale=0.7]{picture/picture1.pdf}}
\subfigure[similarity model]{
\label{Fig.1.sub.2}
\includegraphics[scale=0.55]{picture/picture2.pdf}}
\caption{Two popular structures of building task-specific conversational chatbot}
\label{Fig.1.main}
\end{figure}

@@

\begin{figure}[htbp]
\begin{centering}
\includegraphics[scale=0.48]{picture/picture3}
\par\end{centering}
\caption{\textbf{Network Structure of 2-stage SFC:} a joint system of multi-task for both text classification task and sentence-pair semantic similarity task.}
\label{fig:ver.1.3}
\end{figure}

@@

\begin{figure*}[t]
\begin{centering}
\includegraphics[scale=0.72]{picture/picture4} 
\par\end{centering}
\caption{\textbf{Network Structure of hierarchical SFC:} a joint model of sentence-pair model and text classification model organized in hierarchical structure by completing different tasks at different network layers.}
\label{fig:framework}
\end{figure*}

@@

$\bullet$ \textbf{Full-scale real-world task-specific faq chatbot dataset}

@@

$\bullet$ \textbf{Subset of customer review dataset for extreme classification}

@@

$\bullet$ \textbf{Subset of recently-released intent detection datasets}.

@@

\textbf{Comparison among SFC variants.} First, we turn to compare among the four SFC variants shown in Table \ref{tbe:table2} to analyze the improvement brought by multi-task and hierarchical joint-model structure. We conduct ablation experiments to study the benefits brought by multi-task training by comparing multi-task 2-stage SFC with single-task 2-stage SFC. We can see from Table \ref{tbe:table2} that 2-stage SFC with only task 1 (sentence-pair similarity task) degrades by 0.65 percent on average and 2-stage SFC with only task 2 (top-K classification task) degrades by 0.3 percent on average. The degradation indicates that training with these 2 tasks together in SFC can effectively help the sentence-pair model extract more knowledge from the limited amount of data. Besides, task 2 degrades 0.35 percent less than task 1. This observation suggests that task 2 may play a relatively more important role in multi-task learning, which leads us to put larger weight $\alpha_{K}$ to the loss $\mathcal{L}^{K}$ of task 2 when calculating the total loss $\mathcal{L}$. Also, we can see that the hierachical SFC steadily outperforms 2-stage SFC on 5 diverse datasets by 2 percents on average. This observation supports the idea that the hierachical joint model structure can alleviate the limitation brought by the lack of interaction between similarity model and classification model in 2-stage SFC. Therefore, for the following analysis, we will focus on the comparison between hierachical SFC and the other baseline models.

\textbf{Analysis under few-shot scenarios.} Next, we analyze the overall improvement trend of SFC under various few-shot settings. Our main focus is to study chatbot building in common situation during real-world application where only a few sample sentences are available for each class. The experimental results for the 3 diverse intent detection datasets under few-shot settings indicate that our proposed SFC can achieve +2.47 percents average improvement over one of the most powerful baseline models, which is RoBERTa-large classification model, on all these few-shot settings. Moreover, the improvement in F-score becomes more and more prominent with the decrease of data sample number per class. Our hierarchical SFC can achieve +4.97 percents average improvements over RoBERTa-large classification model in the most extreme few-shot setting with only 5 training samples per class.

\textbf{Fusion of classification and similarity model.} Then, we begin to compare the general performance of SFC with single-task classification and similarity baselines. As for the classification model, hierarchical SFC outperforms BERT-large by +9.81 percents, ALBERT-xxlarge by 4 percents, RoBERTa-large by +2.47 percents, RoBERTa-base by +4.52 percents on average F-score for 5 datasets. In this way, we can see that RoBERTa is the best pre-trained model for the low-resource text classification task we are studying. As for the similarity model, we only choose the strongest pretrained model RoBERTa-large as our baseline, and hierarchical SFC achieves the improvement of 3 percents on average. Moreover, hierarchical SFC's result on 2 datasets without few-shot settings, which are ITG and Amazon-670k, are also 1.24 percents and 2.89 percents higher than Roberta-large classification model. These improvements illustrate that putting classification model and similarity model at different network layers to cooperate with each other is a successful design of hierarchical network structure for joint model, which enables these two models assist each other for better overall results.

\textbf{Settings of hyperparameters.} Finally, we analyze the effect brought by different settings of hyperparameters, see Table \ref{tbe:table3}. We conduct experiments on 5 different settings of hyperparameters, which are the number of candidate class number $K$ and the number of sentence pairs $P$ for each class, on ITG and Amazon-670K datasets to find the most suitable setting of these 2 hyperparameters. We control the total number of sentence pairs sampled for each user query into about the same level of amount(50-60). Through analyzing the experimenal results shown in Table \ref{tbe:table3}, we find that the best setting of hyperparameters might be slightly different for datasets of different data size. The class label number of ITG and Amazon-670k are similar in our experimental setting, which is around 250. However the total amount of data points of ITG is approximately 2 times larger than Amazon-670k. Therefore, the setting of $K=5$ and $P=10$ is good enough for ITG since most of the classes contains over 10 data points. However, for Amazon-670k, the amount of data sample for each class is scarce, thus, we may need more candidate classes to sample enough effective sentence pairs for sentence-pair model layer of SFC. In this way, the setting of $K=10$ and $P=5$ becomes a good choice.
